---
title: "Practical Machine Learning Project"
author: "Kimberly Yuen"
date: "2023-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

##### Background

In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants where they were asked to perform barbell lifts correctly and incorrectly in 5 different ways. I will predict the manner in which they did the exercise - from creating the model, estimating out of sample error and prediction.

## Data Preparation

Load package and read in training and test data.

```{r}
library(caret)
traincsv <- read.csv("pml-training.csv")
testcsv <- read.csv("pml-testing.csv")
```

Remove unnecessary variables, for eg variables that are almost always NA.

```{r}
mostlyNA <- sapply(traincsv, function(x) mean(is.na(x))) > 0.95
train1 <- traincsv[, mostlyNA==F]
```

Remove variables with nearly zero variance.

```{r}
nzv <- nearZeroVar(train1)
train1 <- train1[, -nzv]
```

Remove variables not needed for prediction, the first 5 variables (X, user_name, raw_timestamp_part_1, raw_timestamp_part_2\_,cvtd_timestamp).

```{r}
train1 <- train1[, -(1:5)]
```

Split training data into sub training set and validation set to be able to do cross validation.

```{r}
inTrain <- createDataPartition(y=train1$classe, p=0.7, list=F)
train <- train1[inTrain,]
valid <- train1[-inTrain,]
```

## Build Model

I will test Random Forest model using 3-fold cross validation.

```{r}
fitControl <- trainControl(method="cv", number=3, verboseIter=F)
fit <- train(classe ~ ., data=train, method="rf", trControl=fitControl)
```

Final model has 500 trees and tries 27 variables at each split.

```{r}
fit$finalModel
```

## Results on Validation Set

Prediction on validation set to see Accuracy and Out of Sample error.

```{r}
preds <- predict(fit, newdata=valid)
confusionMatrix(preds,as.factor(valid$classe))
```

**Accuracy is 99.8%, hence out of sample error is 0.2%.** This is very high accuracy so I will use the Random Forests to predict on the test set.

## Prediction

Make prediction of classe on test set.

```{r}
preds2 <- predict(fit, newdata=testcsv)
print(preds2)
```
